---
title: ' Exploratory Data Analysis Assignment2'
author: "FNU Anirudh"
date: "September 16, 2015"
output: html_document
---
Solution 1:-

a)	E [Xi]= ?? and Var[Xi]= ??2
                                 
    _       d
????? (X- ??)------->  N (0, ??^2)

Given ??2 = 1/??2
We can write
     _       d
???n ( X- 1/??) -->???  N (0, 1/??2)


b)	We know that the asymptotic distribution of the sample median is 

     _         d
???n ( X- X0.5) --???>  N (0, 1/4 {f(X0.5) }2)
f (X0.5 )= f (log2/ ??) = ?? e (-?? log 2/??)
            = ??/2
We have , 1/ 4 {  {f(X0.5) }2 = 1 /4 (??2/4d)--->??? N (0,5)
                  -
c) Var (T2)= Var (X/log2)   _
           = (1/ln2)^2 Var (X)
           = 2.081 ??^2   
d) ARE(T1,T2)=    ??^2/n/2.081 ??^2/n
             = 0.481= 48.1%
             
e) Comparing the two statistics T1 and T2, T1 is the better because when we compare the variance of these two methods T1 has a lesser variance, so this would be the preferred statistic method of the given two

f)

source(lvalprogs.r)
x <- rexp(1000, 1)
lval(x)

  Depth  Lower  Upper    Mid Spread pseudo-s
M 500.5 0.6932 0.6932 0.6932 0.0000   0.0000
F 250.5 0.2929 1.3877 0.8403 1.0948   0.8116
E 125.5 0.1291 2.0037 1.0664 1.8746   0.8148
D  63.0 0.0617 2.6406 1.3511 2.5790   0.8405
C  32.0 0.0289 3.3680 1.6985 3.3391   0.8963
B  16.5 0.0149 4.3889 2.2019 4.3741   1.0154
A   8.5 0.0083 4.6239 2.3161 4.6157   0.9546
Z   4.5 0.0050 5.2925 2.6488 5.2874   0.9938
Y   2.5 0.0009 5.6520 2.8264 5.6511   0.9792
X   1.5 0.0006 5.8699 2.9352 5.8692   0.9475
W   1.0 0.0004 5.9646 2.9825 5.9642   0.9044

             
Data is skewed to the right.


Solution 2:-

```{r,eval=TRUE}
library(aplpack)

# Letter Value
lval <- function(x) {
  #tag <- c("M ","F ","E ","D ","C ","B ","A ","Z ","Y ","X ","W ","V","U","T",
  # "S","R","Q","P","O","N")
  # gau <- abs(qnorm(c(.25,.125,1/16,1/32,1/64,1/128,1/256,1/512,1/1024,1/2048,
  #    1/4096, 1/8192, 1/16384, 1/32768, 1/65536)))
  tag <- c("M",LETTERS[6:1],LETTERS[26:14])
  
  gau <- abs(qnorm(1/2^(2:20)))
  
  # col 1 = depth; 2 = lower; 3 = upper; 4 = mid; 5 = spread; 6 = pseudo-s
  
  y <- sort(x[!is.na(x)])
  n <- length(y)
  m <- ceiling(log(n)/log(2)) + 1
  depth    <- rep(0,m)
  depth[1] <- (1 + n)/2
  
  for (j in 2:m) {depth[j] <- (1 + floor(depth[j-1]))/2 }
  
  ndepth <- n+1 - depth
  out <- matrix(0, m, 6)
  dimnames(out) <- list(tag[1:m],
                        c("Depth", "Lower","Upper", "Mid", "Spread","pseudo-s"))
  out[1,2:3] <- median(y)
  out[,1] <- depth
  
  
  for (k in 2:m) {
    out[k,2] <- ifelse(depth[k] - round(depth[k]) == 0,
                       y[depth[k]], (y[depth[k]-.5]+y[depth[k]+.5])/2 )
    out[k,3] <- ifelse(ndepth[k] - round(ndepth[k]) == 0,
                       y[ndepth[k]], (y[ndepth[k]-.5]+y[ndepth[k]+.5])/2 )
  }
  
  out[1:m,4] <- (out[1:m,2] + out[1:m,3])/2
  out[2:m,5] <- out[2:m,3] - out[2:m,2]
  out[2:m,6] <- out[2:m,5]/(2*gau[1:(m-1)])
  round(out,4)
}



A = c(19.50, 16.72, 20.92, 16.42, 21.22, 15.40, 20.68, 14.55, 20.23
      , 15.11, 20.95, 16.68, 14.67, 16.50, 22.15 , 20.14, 18.33, 14.20
      ,11.61,22.24,18.75,14.22,15.03,22.07,13.34,12.73,19.23
      ,19.74,19.74,20.60,19.29,18.22,23.65,17.44,13.07,19.00
      ,18.44, 17.25, 19.19, 12.77,14.10,16.69,16.92,21.92,20.84
      ,18.43, 19.54, 23.61, 21.40, 28.34, 20.43, 20.43, 15.58,16.58
      , 22.44, 14.59, 18.70, 16.79, 14.12, 13.67, 15.94, 24.04 , 15.42
      , 16.26, 17.74, 12.37 , 16.87, 16.28, 17.97, 19.56, 13.56, 16.13
      , 18.20, 17.79, 19.38, 20.47, 16.75, 16.69, 15.93, 14.73, 17.83
      , 19.78, 15.78, 16.17, 17.18, 13.90, 15.33, 16.10, 12.03,17.92
      , 23.56, 11.35, 19.10, 12.91, 18.32, 19.24, 11.57, 14.33, 13.60
      , 13.12, 11.19, 14.33, 16.91, 13.03, 17.32, 10.70, 12.56, 16.04)

# Plot Time series
b= ts(A, frequency = 12, start = c(1997, 1))
plot(b)

# Stem and Leaf
stem.leaf(b)


lval(b)

# QQ Plot
qqnorm(b)
qqline(b)

```

From stem and leaf we can see that data is skewed to the left.
d)  NotNormally distributed
e) Yes, there is outlier.

Solution 3:- 

a) Single Batch n=120

0.4 + 0.007* n= 0.4+ 0.007 * 120= 1.24

b) Two batches n=60

a= 0.4+0.007*n= 0.4+0.42=0.82

Similarly,b= 0.82

Total outside values= a+b= 0.82+0.82= 1.64

c) n=40

a= 0.4+ 0.007* 40= 0.68

n=30

b= 0.4+ 0.007* 30= 0.61

n=20

c= 0.4+0.007* 20= 0.54

n=10 

c= 0.4+0.007* 10= 0.47

n=5 

c= 0.4+ 0.007*5= 0.435

Total outside value= 0.68 + 0.61 + 0.54 + 0.47 + 4* 0.435 = 4.04

Solution 4

1)


```{r, eval=TRUE}
library(cluster)
vote=votes.repub
vote_calc= function(vote,lower,upper)
{
  colours=c("blue","black","red","deeppink","red","purple","green","brown",
            "orange","violet")
  par(xpd=NA,oma=c(0,0,0,6))
  for(i in lower:upper)
  {
    vote_t=vote[i,]
    vote_t=as.data.frame(t(vote_t))
    vote_ts=ts(vote_t,frequency=0.25,start=1856)
    if(i==lower)
    {
      plot(vote_ts,col=colours[1],ylim=range(0,100),xlim=range(1856,1976),
           ylab=paste(lower,"-",upper,"states"))
    }
    else
      lines(vote_ts,col=colours[i%%10])
  }
  legend(1982,150,legend=row.names(vote[lower:upper,]),
         fill=colours,title="States")
  segments(1852,50,1980,50,col="grey65",lty=2)
}
vote_calc(vote,1,10)
vote_calc(vote,11,20)
vote_calc(vote,21,30)
vote_calc(vote,31,40)
vote_calc(vote,41,50)

```

2)

```{r, eval=TRUE}
library(cluster)
vote = votes.repub
par(mfrow = c(3, 2))
vote_calc = function(vote, reg, name)
{
  
  colours=c("blue","black","red","deeppink","red","purple","green","brown",
            "orange","violet")
  par(mar=c(4,5,0,2))
  j = 1
  for(i in reg)
  {
    vote_t = vote[i,]
    vote_t = as.data.frame(t(vote_t))
    vote_ts = ts(vote_t, frequency = 0.25, start = 1856)
    if(i==reg[1])
    {
      plot(vote_ts, col = colours[1], ylim = range(0,100),
           ylab = name)
    }
    else
      j = j + 1
    lines(vote_ts, col = colours[j])
  }
  segments(1852,50,1980,50, col = "grey65", lty = 2)
}

northeast = c("Connecticut","Delaware","Maine", "Massachusetts","New Hampshire","New Jersey",
              "New York","Pennsylvania","Rhode Island","Vermont")
east_central = c("Kentucky","Maryland","North Carolina","South Carolina","Tennessee",
                 "Virginia","West Virginia")
south = c("Alabama", "Arkansas", "Florida", "Georgia","Louisiana","Mississippi","Oklahoma","Texas")
midwest = c("Illinois","Indiana","Iowa","Kansas","Michigan","Minnesota","Missouri","Nebraska",
            "Ohio","Wisconsin")
rockies = c("Colorado","Idaho","Montana","North Dakota","South Dakota","Utah","Wyoming")
west = c("Alaska","Arizona","California","Hawaii","Nevada","New Mexico","Oregon","Washington")

vote_calc(vote, northeast, "Northeast")
vote_calc(vote, east_central, "Mid-Atlantic / East-Central")
vote_calc(vote, south, "South")
vote_calc(vote, midwest, "Midwest")
vote_calc(vote, rockies, "Rockies")
vote_calc(vote, west, "Western")

```

Republicans recieved less votes from South.

3).

```{r,eval=TRUE}
library(cluster)
vote = votes.repub
par(mfrow = c(3, 2))
vote_calc = function(vote, reg, name)
{
  
  colours=c("blue","black","red","deeppink","red","purple","green","brown",
            "orange","violet")
  par(mar=c(4,5,0,2))
  j = 1
  for(i in reg)
  {
    vote_t = vote[i,]
    vote_t = as.data.frame(t(vote_t))
    vote_ts = ts(vote_t, frequency = 0.25, start = 1856)
    if(i==reg[1])
    {
      plot(vote_ts, col = colours[1], ylim = range(0,100),
           ylab = name)
    }
    else
      j = j + 1
    lines(vote_ts, col = colours[j])
  }
  segments(1852,50,1980,50, col = "grey65", lty = 2)
}

northeast = c("Connecticut","Delaware","Maine", "Massachusetts","New Hampshire","New Jersey",
              "New York","Pennsylvania","Rhode Island","Vermont")
east_central = c("Kentucky","Maryland","North Carolina","South Carolina","Tennessee",
                 "Virginia","West Virginia")
south = c("Alabama", "Arkansas", "Florida", "Georgia","Louisiana","Mississippi","Oklahoma","Texas")
midwest = c("Illinois","Indiana","Iowa","Kansas","Michigan","Minnesota","Missouri","Nebraska",
            "Ohio","Wisconsin")
rockies = c("Colorado","Idaho","Montana","North Dakota","South Dakota","Utah","Wyoming")
west = c("Alaska","Arizona","California","Hawaii","Nevada","New Mexico","Oregon","Washington")

# Box Plot for All States
par(mfrow=c(1,1))
boxplot(t(votes.repub), main ="All States")
qqnorm(t(votes.repub), main ="All States")

par(mfrow=c(3,2))
# Box Plot for Northwest
boxplot(t(votes.repub[northeast,]))

# Box Plot for East central
boxplot(t(votes.repub[east_central,]))

#Box Plot for South
boxplot(t(votes.repub[south,]))

#Box Plot for Midwest
boxplot(t(votes.repub[midwest,]))

# Box Plot for Rockies
boxplot(t(votes.repub[rockies,]))

# Box Plot for West
boxplot(t(votes.repub[west,]))


par(mfrow=c(3,2))
qqnorm(t(votes.repub[northeast,]), main ="North East")
qqnorm(t(votes.repub[east_central,]), main = "East central")
qqnorm(t(votes.repub[south,]), main="South")
qqnorm(t(votes.repub[midwest,]), main="Midwest")
qqnorm(t(votes.repub[rockies,]), main="Rockies")
qqnorm(t(votes.repub[west,]), main="West")


```



